---
title: ' '
output:
  html_document:
    highlight: tango
    css: ~/.s/style.css
---

<style>
.sourceCode
  { background-color: initial; }
</style>
\newcommand{\EE}{\mathbb{E}}
\newcommand{\PP}{\mathbb{P}}

# In-term I, M362M Spring 2024


Submit your solution in pdf (knitted from Rmd) just like you would submit your HW. Additionally, **submit the Rmd file you used to knit the pdf**.   Submitting only the pdf file will result in a 50\% penalty.

If you have any questions, please post them on Ed Discussion.

In all problems, choose the number `nsim` of simulations as large as possible without the code taking more than a few seconds. 


## Problem 1.

```{r child="~/.s/problems/00_R/Sim-joint-cond-3_prb.Rmd"}
```

<!-- *Solution* -->

<!-- ```{r child="~/.s/problems/00_R/Sim-joint-cond-3_sol.Rmd"} -->
<!-- ``` -->

## Problem 2.

```{r child="~/.s/problems/00_R/Money-growth_prb.Rmd"}
```

<!-- *Solution* -->

<!-- ```{r child="~/.s/problems/00_R/Money-growth_sol.Rmd"} -->
<!-- ``` -->


