
```{r child="mydefs.Rmd"}
```
# Random Walks
<div style="counter-reset: thechapter 3;"> </div>


## What are stochastic processes?

A **stochastic process** is a sequence - finite or infinite - of random
variables. We usually write $\{X_n\}_{n\in\N_0}$  or $\{X_n\}_{0\leq n \leq T}$,
depending on whether we are talking about an infinite or a finite sequence. The
number $T\in \N_0$ is called the **time horizon**, and we sometimes set
$T=+\infty$ when the sequence is infinite. The index $n$ is often interpreted
as *time*, so that a stochastic process can be thought of as a model of a random
process evolving in time. The initial value of the index $n$ is often normalized
to $0$, even though other values may be used. This it usually very clear from
the context.

It is important that all the random variables $X_0, X_1,\dots$ "live" on the
same sample space $\Omega$. This way, we can talk about the notion of a
**trajectory** or a **sample path** of a stochastic process: it is, simply, the
sequence of numbers $$X_0(\omega), X_1(\omega), \dots$$ but with $\omega\in
\Omega$ considered "fixed". In other words, we can think of a stochastic
process as a random variable whose value is not a number, but  sequence of
numbers. This will become much clearer once we introduce enough examples.

## The Simple Symmetric Random Walk

A stochastic process $\{X\}_{n\in\N_0}$ is said to be a 
**simple symmetric random walk (SSRW)** if

1.  $X_0=0$,

2.  the random variables $\delta_1 = X_1-X_0$, 
$\delta_2 = X_2 - X_1$, ..., called the **steps** of the random walk, are independent

3.  each $\delta_n$ has a **coin-toss distribution**, i.e., its distribution 
is given by $$\PP[ \delta_n = 1] = \PP[ \delta_n=-1] = \tfrac{1}{2} \text{ for each } n.$$

Some comments:

 - This definition captures the main features of an
    idealized notion of a particle that gets shoved, randomly, in one of
    two possible directions, over and over. In other words, these "shoves"
    force the particle to take a step, and steps are 
    modeled by the random variables
    variables $\delta_1,\delta_2, \dots$. The position of the
    particle after $n$ steps is $X_n$; indeed,
    $$X_n = \delta_1 + \delta_2 + \dots + \delta_n \text{ for }n\in \N.$$ It
    is important to assume that any two steps are independent of each
    other -  the most important properties of random walks depend on this
    in a critical way.

- Sometimes, we only need a finite number of steps of a random walk,
    so we only care about the random variables $X_0, X_1,\dots, X_T$. This
    stochastic process (now with a finite time horizon $T$) will also be
    called a random walk. If we want to stress that the horizon is not infinite, 
    we sometimes call it the **finite-horizon random walk**. Whether $T$ is 
    finite or infinite is usually 
    clear from the context.
    
- The starting point $X_0=0$ is just a normalization. Sometimes we
    need more flexibility and allow our process to start at $X_0=x$ for
    some $x\in \N$. To stress that fact, we talk about the random walk **starting at $x$**.
    If no starting point is mentioned, you should assume $X_0=0$.

-  We will talk about **biased** (or **asymmetric**) random walks a bit later. 
    The only
    difference will be that the probabilities of each $\delta_n$ taking
    values $1$ or $-1$ will be $p\in (0,1)$ and $1-p$,  and not necessarily $\tot$, 
    The probability $p$ cannot change 
    from step to step and the steps $\delta_1, \delta_2, \dots$ will continue to be 
    independent from each other.
    
- The word simple in its name refers to the fact that distribution of every 
    step is a coin toss. You can easily imagine a more complicated mechanism that 
    would govern each step. For example, not only the direction, but also the 
    size of the step could be random. In fact, any distribution you can think
    of can be used as a step distribution of a random walk. Unfortunately, we will 
    have very little to say about such, general, random walks in these notes.

## How to simulate random walks

In addition to being quite simple conceptually, random walks are also easy 
to simulate. The fact that the steps $\delta_n = X_n - X_{n-1}$ are independent
coin tosses immediately suggests a feasible strategy: 
simulate $T$ independent coin tosses first, and then define each 
$X_n$ as the sum of the first $n$ tosses. 

Before we implement this idea in R, let us agree on a few conventions which 
we will use whenever we simulate a stochastic process:

- the result of each simulation is a `data.frame` object
-  its columns will be the random variables $X_0$, $X_1$, \dots It is a good
idea to name your columns `X0`, `X1`, `X2`, etc.
- each row will represent one "draw"˜
˜µ
This is best achieved by the following two-stage approach in R:

a) write a function which will simulate a single trajectory 
of your process, If
your process comes with parameters, it is a good idea to include them
as arguments to this function. 

b) use the function `replicate` to stack together many such simulations and 
convert the result to a `data.frame`. Don't 
forget to transpose first (use the function `t`) because `replicate` works column by column, 
and not row by row. 



Let's implement this in the case of a simple random walk. Of course, it is
impossible to simulate a random walk on an infinite horizon ($T=\infty$) 
so we must restrict to finite-horizon random walks^[It is somewhat unfortunate that the standard notation for the time horizon, namely $T$, coincides with a
shortcut `T` for `TRUE` in R. Our example still works fine because this shortcut is used only if there is no variable named `T`.]. The function `cumsum` which
produces partial sums of its input comes in very handy.

```{r}
single_trajectory = function(T, p=0.5) {
    delta = sample(c(-1,1), size=T, replace=TRUE, prob=c(1-p,p))
    x = cumsum(delta)
    return(x)
}
```

Next, we run the same function `nsim` times and record the results. It is a
lucky break that the default names given to columns are `X1`, `X2`, ... so we
don't have to rename them. We do have to add the zero-th column $X_0=0$ because, 
formally speaking, the "random variable" $X_0=0$ is a part of the stochastic
process. This needs to be done before other columns are added to maintain the
proper order of columns, which is important when you want to plot trajectories.

```{r echo=-2, tidy=FALSE }
simulate_walk = function(nsim, T, p = 0.5) {
  return(
    data.frame(
      X0 = 0,
      t(replicate(nsim, single_trajectory(T, p)))
    ))
}
set.seed(2011)
walk = simulate_walk(nsim = 10000, T = 500)
```


## Two ways of looking at a stochastic proceses

Now that we have the data frame `walk`, we can explore in at least two qualitatively different ways:

### Column-wise (distributionally)

Here we focus on individual random variables (column) or pairs, triplets, etc. of random variables and study their (joint) distributions. For example, we can plot histograms of the random variables $X_5, X_8, X_{30}$ or $X_{500}$:

<center>

```{r echo=FALSE, fig.width=11, out.width="100%", chache=TRUE}
library(latex2exp)
par(mfrow=c(2,2))
hist(
    walk$X5,
    prob = TRUE,
    breaks = seq(-5.5, 5.5, by = 1),
    xaxt = "n",
    xlab = "",
    ylim=c(0,0.31),
    main = TeX('Histogram of $X_5$')
)
axis(side=1, at=seq(-5,5, 2), labels=seq(-5,5,2) )

hist(
    walk$X8,
    prob = TRUE,
    breaks = seq(-8.5, 8.5, by = 1),
    main = TeX('Histogram of $X_8$'),
    xlab="",
    xaxt = "n",
    ylim = c(0,0.31)
)
axis(side=1, at=seq(-8,8, 2), labels=seq(-8,8,2) )

hist(
    walk$X30,
    prob = TRUE,
    xaxt = "n",
    xlab = "",
    xlim=c(-30,30),
    main = TeX('Histogram of $X_{30}$')
)
axis(side=1, at=seq(-30,30, 5), labels=seq(-30,30,5) )

hist(
    walk$X500,
    prob = TRUE,
    xlim=c(-500,500),
    main = TeX('Histogram of $X_{500}$'),
    xlab="",
    xaxt = "n"
)

axis(side=1, at=seq(-500,500, 50), labels=seq(-500,500,50) )

```

</center>

We can also use various (graphical or not) devices to understand joint distributions of pairs of random variables:

<center>
```{r echo=FALSE}
library(ggplot2)
tf=data.frame(table(factor(walk$X5, levels=-5:5), factor(walk$X8, levels=-8:8)))
tf$Var1=as.numeric(tf$Var1)-6
tf$Var2=as.numeric(tf$Var2)-9

ggplot(data=tf, aes(xmin=Var2-0.5, xmax = Var2+0.5, ymin = Var1 - 0.5, ymax = Var1 + 0.5))+
    geom_rect(aes(fill=Freq), color="grey50")+
    coord_fixed()+
    scale_fill_gradient(low="white", high="brown4", trans="sqrt")+
    theme_classic() +
    theme(legend.position="none")+
    scale_x_continuous(breaks = seq(-8,8,by=1))+
     scale_y_continuous(breaks = seq(-5,5,by=1))+
     xlab(TeX("$X_8$"))+
    ylab(TeX("$X_5$"))
```
</center>

### Row-wise (trajectorially or path-wise)

If we focus on what is going on in a given row of `walk`, we are going to see 
a different cross-section of our stochastic process. This way we are fixing the
state of the world $\omega$ (represented by a row of `walk`), i.e., the particular
realization of our process, but
varying the time parameter. A typical picture associated to a trajectory 
of a random walk is the following



```{r warning=FALSE, echo=FALSE, message=FALSE, fig.align="center"}
library(tidyverse)
library(gridExtra)

set.seed(1141)

T=10;
x = c(0,single_trajectory(T))
t = 0:T
walk = data.frame( t,x)
plot1 = ggplot(data=walk, aes(x=t, y=x)) + 
    geom_line()+geom_point(size=1)+
    ggtitle("A trajectory of a random walk, T=10")+
    theme_bw()

T=100;
x = c(0, single_trajectory(T))
t = 0:T
walk = data.frame( t,x)
plot2 = ggplot(data=walk, aes(x=t, y=x)) + geom_point(size=0.1)+geom_line()+ggtitle("A trajectory of a random walk, T=100")+theme_bw()

grid.arrange(plot1, plot2, nrow=2)
```

You can try to combine the two approaches (if you must) and plot several trajectories on the same plot. 
While this produces pretty pictures (and has one or two genuine applications), it usually leads to a 
sensory overload. Note that the trajectories below are jittered a bit. That means that the positions of the points are randomly shifted by a small amount. This allows us to see features of the plot that would otherwise be hidden because of the overlap.

<center>

```{r warning=FALSE, message=FALSE, echo=FALSE, fig.height=3}
library(tidyverse)
T=15
nsim=10

walk_long = simulate_walk(nsim, T=T) %>% 
    setNames(as.character(0:T)) %>% # they are X0, X1, etc. originally
    rownames_to_column(var="omega") %>% # row number is omega
    pivot_longer(-omega,
                 names_to = "t", 
                 values_to = "X") %>% # pivot to a function of (t,omega)
    mutate(t = as.numeric(t))

walk_jittered=walk_long %>% 
    mutate(t = t+rnorm(length(t),0,0.1))

ggplot(walk_jittered, aes(x=t, y=X, group = omega, color=omega))+geom_line(size=1)+theme_bw()+xlab("")+ylab("")+
  theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank())+
          coord_fixed()
```

</center>


## The path space

The row-wise (or path-wise or trajectory-wise) view of the random walk described above illustrates a 
very important point: the random walk (and random processes in general) can be
seen as random "variable" whose values are not merely numbers; they are sequences of numbers (trajectories). 
In other words, a random process is simply a "random trajectory". 
We can simulate this random trajectory as we did above, but simulating the steps and adding them up, but we could also 
take a different approach. We could build the set of
all possible trajectories, and then pick a random trajectory out of it. 

For a random walk on a finite horizon $T$, a trajectory is
simply a sequence of natural numbers starting from $0$. Different
realizations of the coin-tosses $\delta_n$ will lead to different
trajectories, but not every sequence of natural numbers corresponds to a
trajectory. For example $(0,3,4,5)$ is not possible because the increments of the random walk can only take values $1$ or $-1$. In fact, a finite
sequence $(x_0, x_1, \dots, x_T)$ is a (possible) sample path of a
random walk if and only if $x_0=0$ and $x_{k}-x_{k-1} \in \{-1,1\}$
for each $k$. For example, when $T=3$, there are $8$ possible trajectories:
$$ \begin{align} \Omega = \{ 
&(0,1,2,3), (0,1,2,1),(0,1,0,2), (0,1,0,-1), \\
& (0,-1,-2,-3), (0,-1,-2,-1), (0,-1,0,-2), (0,-1,0,1)\}
\end{align}$$
When you (mentally) picture them, think of  their graphs:

<center>

```{r warning=FALSE, echo=FALSE}
library(ggplot2)
library(gridExtra)
paths =list(c(0,1,2,3), c(0,1,2,1), c(0,1,0,1), c(0,1,0,-1),
c(0,-1,0,1), c(0,-1,0,-1), c(0,-1,-2,-1), c(0,-1,-2,-3))

pl=list(8)
i=1
for (path in paths) {
  df = data.frame(t=0:3, path=path)
  pl[[i]] = ggplot(data=df, aes(x=t, y=path)) + 
    geom_line()+geom_point(size=1)+
    theme_bw()+coord_fixed(ratio=0.5)+xlab("")+ylab("")+
    ylim(-3,3)
  i=i+1
}
do.call("grid.arrange", c(pl, ncol=4))
```
</center>

Each trajectory corresponds to a particular combinations of the values of the 
increments $(\delta_1,\dots, \delta_T)$, each such combination happens with probability $2^{-T}$. 
This means that any two trajectories are equally likely. That is convenient, because
this puts uniform probability on the collection of trajectories. We are now ready to 
implement our simulation procedure in R; let us write the function `single_trajectory` using this
approach and use it to simulate a few trajectories. We assume that a function `all_paths(T)` which returns
a list of all possible paths with horizon $T$ has already been implemented (more info about a possible 
implementation in R is given in a problem below):

```{r echo=c(5,6,7,8,9), tidy=FALSE}
choice_to_path = function(comb,T) {
  increments = rep(-1,T)
  increments[comb]=1
  path = cumsum(increments)
  return(path)
}

all_paths = function (T) {
  Omega = list(2^T)
  index=1;
  for (i in 0:T) {
    choices = combn(T, i, simplify=FALSE)
    for (choice in choices) {
      Omega[[index]] =  choice_to_path(choice,T)
      index=index+1
    }
  }
  return(Omega)
}

T=5
Omega = all_paths(T)
single_trajectory = function() {
  return(unlist(sample(size=1,Omega)))
}

simulate_walk = function(nsim, p=0.5) {
   return(data.frame(
            X0=0, 
            t(replicate(nsim, single_trajectory()))
        ))
}
```


## The distribution of $X_n$

Building a path space is not simply an exercise in abstraction. Here is how we can use is to
understand the distribution of the position of the random walk:

<div class="example">
Let $X$ be a simple symmetric random walk with time horizon $T=5$. What is the probability that $X_{5}=1$?
</div>
<div class="solution">
Let $\Omega$ be the path space, i.e., the set of all possible trajectories of length $5$ - there are $2^{5}=32$ of them. The probability that $X_{5}=1$ is the probability that a randomly picked path from $\Omega$ will take the value $1$ at $n=5$. Since all paths are equally likely, we need to *count*
the number of paths with value $1$ at $n=5$ and then divide by the total number of paths, i.e., $32$.

So, how many paths are there that take value $1$ at $n=5$? Each path is built out of steps of absolute value $1$. Some of them go up (call them up-steps) and some of them go down (down-steps). A moment's though reveals that the only way to reach $1$ in $5$ steps is if you have exactly $3$ up-steps and $2$ down-steps. Conversely, any path that has $3$ up-steps and $2$ down-steps ends at $1$. 

This realization transforms the problem into the following: how many paths are there with exactly $3$ up-steps (note that we don't have to specify that there are $2$ down-steps - it will happen automatically). The only difference between different paths with exactly $3$ up-steps is the position of these up-steps. In some of them the up-steps happen right at the start, in some at the very end, and in some they are scattered around. Each path with $3$ up-steps is uniquely determined by the list of positions of those up-steps, i.e., with a size-$3$ subset of $\{1,2,3,4,5\}$. This is not a surprise at all, since each path is build out of increments, and positions of positive increments clearly determine values of all increments.

The problem has now become purely mathematical: how many size-$3$ subsets of $\{1,2,3,4,5\}$ are there? The answer comes in the form of a *binomial coefficient* $\binom{5}{3}$ whose value is $10$ - there are exactly ten ways to pick three positions out of five. Therefore,
$$ \PP[ X_{5} = 1] = 10 \times 2^{-5} = \frac{5}{16}.$$
</div>

Can we do this in general?

<div class="example">
Let $X$ be a simple symmetric random walk with time horizon $T$. What is the probability that $X_{n}=k$?
</div>
<div class="solution">
The reasoning from the last example still applies. A trajectory with $u$ up-steps and $d$ down-steps 
will end at $u-d$, so we must have $u-d=k$. On the other hand $u+d=n$ since all steps that are not up-steps are necessarily down-steps. This gives as a simple linear system with two equations and two unknowns which solves to $u = (n+k)/2$, $d=(n-k)/2$. Note the $n$ and $k$ must have the same parity for this solution to be meaningful. Also, $k$ must be between $-n$ and $n$. 

Having figured out how many up-steps is necessary to reach $k$, all we need to do is count the number of trajectories with that many up-steps. Like before, we can do that by counting the number of 
ways we can choose their position among $n$ steps, and, like before, the answer is the binomial coefficient $\binom{n}{u}$ where $u=(n+k)/2$. Dividing by the total number of trajectories gives us the final answer:
$$ \PP[ X_n = k ] = \binom{n}{ (n+k)/2} 2^{-n},$$
for all $k$ between $-n$ and $n$ with same parity as $n$. For all other $k$, the probability is $0$.
</div>

The binomial coefficient and the $n$-th power suggest that the distribution of $X_n$ might have something
to do with the binomial distribution. It is clearly not the binomial, since it 
can take negative values, but it is related. To figure out what is going on, let us first remember what the binomial distribution is all about. Formally, it is a discrete distribution with two parameters $n\in\N$ and $p\in (0,1)$. Its support is $\set{0,1,2,\dots, n}$ and the distribution is given by the following table, where $q=1-p$

<center>
<div style="width: 80%">
0 | 1 | 2 | ... | k | ... |n
:---:|:---:|:---:|:---:|:---:|:---:|:---:|
 $\binom{n}{0} q^n$ | $\binom{n}{1} p q^{n-1}$ | $\binom{n}{2} p^2 q^{n-2}$ | ... | $\binom{n}{k} p^k q^{n-k}$ | ... | $\binom{n}{n} p^n$
</div>
</center>

The binomial distribution is best understood, however, when it is expressed as a "number of successes". 
More precisely, 

> If $B_1,B_2,\dots, B_n$ are $n$ *independent* Bernoulli random variables with 
> *the same* parameter $p$, then their sum $B_1+\dots+B_n$ has a binomial distribution with 
> parameters $n$ and $p$.

We think of $B_1, \dots, B_n$ as indicator random variables of "successes" in $n$ independent
"experiments" each of which "succeeds" with probability $p$. A canonical example is tossing a biased coin
$n$ times and counting the number of "heads". 

We know that the position $X_n$ at time $n$ of the random walk admits the representation
$$ X_n = \delta_1+\delta_2+\dots+\delta_n,$$
just like the binomial random variable. The distribution of $\delta_k$ is not Bernoulli, though, since it takes the values $-1$ and $1$, and not $0,1$. This is easily fixed by applying the linear transformation $x\mapsto \ot (x+1)$; indeed $( -1 +1)/2 = 0$ and $( 1 + 1) / 2 =1$, and, so, 
$$ \ot (\delta_k+1)\text{ is a Bernoulli random variable with parameter } p=\frac{1}{2}.$$
Consequently, if we add all $B_k = \tot(1+\delta_k)$ and remember our discussion from above we get the following statement

> In a simple symmetric random walk the random variable $\frac{1}{2} (n + X_n)$ has 
> the binomial distribution with parameters $n$ and $p=1/2$, for each $n$.

Can you use that fact to rederive the distribution of $X_n$? 

## Biased random walks

If the steps of the random walk preferred one direction to the other,
the definition would need to be tweaked a little bit and the word "symmetric" in the name gets replaced by "biased" (or "asymmetric"):

A stochastic process $\{X\}_{n\in\N_0}$ is said to be a 
**simple biased random walk with parameter $p\in (0,1)$ if

1.  $X_0=0$,

2.  the random variables $\delta_1 = X_1-X_0$, 
$\delta_2 = X_2 - X_1$, ..., called the **steps** of the random walk, are independent and

3.  each $\delta_n$ has a **biased coin-toss distribution**, i.e., its distribution 
is given by
    $$\PP[ \delta_n = 1] = p \text{ and } \PP[ \delta_n=-1] = 1-p \text{ for each } n.$$

As far as the distribution of $X_n$ is concerned, we don't expect 
it to be the same as in the symmetric case. After all, the 
biased random walk (think $p=0.999$) will prefer one direction over 
the other. Our trick 
with writing $\ot(n+X_n)$ as a sum of Bernoulli random variables still works. 
We just have to remember 
that $p$ is not $\ot$ anymore to conclude that $\tot (X_n + n)$ has the binomial 
distribution with parameters $n$ and $p$; if we put $u = (n+k)/2$ we get
\begin{align}
\PP[ X_n = k] &= \PP[ \tot (X_n+n) = u] = \binom{n}{u} p^u q^{n-u}\\ & = \binom{n}{\ot (n+k)} p^{\ot (n+k)} q^{\ot (n-k)}. 
\end{align}
Note that be binomial coefficient stays the same as in the symmetric case, 
but the factor $2^{-n} = (1/2)^{\ot (n+k)} (1/2)^{\ot(n-k)}$ becomes 
$p^{\ot (n+k)} q^{\ot (n-k)}$.

Can we reuse the sample space $\Omega$ to build a
biased random walk? Yes, we can, but we  need to
assign possibly different probabilities to individuals. Indeed, if
$p=0.99$, the probability that all the increments $\delta$ of a $10$-step 
random walk take the
value $+1$ is $(0.99)^{10} \approx 0.90$. This is much larger 
than the probability that all steps take the value $-1$, which is $(0.01)^{10}= 10^{-20}$. 

In general, the probability that a particular path is picked out of $\Omega$
will depend on the number of up-steps and down-steps; more precisely it equals
$p^u q^{n-u}$ where $u$ is the number of up-steps. The interesting thing is that
the number of up-steps $u$ depends only on the final position $x_n$ of the path; indeed $u =
\ot (n+x_n)$. This way, all paths of length $T=5$ that end up at $1$ get the
same probability of being chosen, namely $p^3 q^2$. Let us use the awful
seizure-inducing graph with multiple paths for good, and adjust the each path
according to its probability; some jitter has been added to deal with overlap.
The lighter-colored paths are less likely to happen then the darker-colored paths.

<center>

```{r warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(viridis)

single_trajectory = function(T, p=0.5) {
    delta = sample(c(-1,1), size=T, replace=TRUE, prob=c(1-p,p))
    x = cumsum(delta)
    return(x)
}

simulate_walk = function(nsim, T, p=0.5) {
   return(data.frame(
            X0=0, 
            t(replicate(nsim, single_trajectory(T, p)))
        ))
}


nsim = 30
T=30
beta = 0
lambda = -0.001
set.seed(16891)
last_name = str_c("X",T)
walk = simulate_walk(nsim, T) %>% 
    setNames(as.character(0:T)) %>% # they are X0, X1, etc. originally
    rownames_to_column(var="omega") 

walk$weight = ordered(walk[,T+2])

walk_long = walk %>% # row number is omega
    pivot_longer(-c(omega,weight),
                 names_to = "t", 
                 values_to = "X") %>% # pivot to a function of (t,omega)
    mutate(t = as.numeric(t))

walk_jittered = walk_long %>% 
    mutate(t = t+rnorm(length(t),0,0.1))

ggplot(walk_jittered, aes(x=t, y=X, group = omega, color=weight))+geom_line(size=1)+theme_bw()+xlab("")+ylab("")+
  theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank())+
          coord_fixed()+scale_color_brewer(palette="Blues")
```

</center>




## Additional problems for Chapter 3

<div class="problem">
```{r child="problems/01_Random_Walks/walk6_prb.Rmd"}
```
</div>
<div class="solution">
```{r child="problems/01_Random_Walks/walk6_sol.Rmd"}
```
</div>


<div class="problem">
```{r child="problems/01_Random_Walks/simple-1_prb.Rmd"}
```
</div>
<div class="solution">
```{r child="problems/01_Random_Walks/simple-1_sol.Rmd"}
```
</div>

<div class="problem">
```{r child="problems/01_Random_Walks/simple-1b_prb.Rmd"}
```
</div>
<div class="solution">
```{r child="problems/01_Random_Walks/simple-1b_sol.Rmd"}
```
</div>


<!--
  same_value_three_points
-->
<div class="problem">
```{r child="problems/01_Random_Walks/same_value_three_points_prb.Rmd"}
```
</div>
<div class="solution">
```{r child="problems/01_Random_Walks/same_value_three_points_sol.Rmd"}
```
</div>

<!--
  R-impl-all paths
-->
<div class="problemec">
```{r child="problems/01_Random_Walks/R-impl-all-paths_prb.Rmd"}
```
</div>
<div class="solution">
```{r child="problems/01_Random_Walks/R-impl-all-paths_sol.Rmd"}
```
</div>


<!--
  var_cov_corr
  ------------------------------------------------
-->
<div class="problem">
```{r child="problems/01_Random_Walks/var_cov_corr_prb.Rmd"}
```
</div>
<div class="solution">
```{r child="problems/01_Random_Walks/var_cov_corr_sol.Rmd"}
```
</div>

<!--
  area_under_walk
  ------------------------------
-->
<div class="problem">
```{r child="problems/01_Random_Walks/area_under_walk_prb.Rmd"}
```
</div>
<div class="solution">
```{r child="problems/01_Random_Walks/area_under_walk_sol.Rmd"}
```
</div>

⬇︎ In case you were wondering, the text below belongs to footnotes from somewhere high above.⬇︎
