---
knit: |
  (function(input, ...) {
    rmarkdown::render(
      input,
      output_file = paste0(
        xfun::sans_ext(input), '-', Sys.Date(), '.html'
      ),
      envir = globalenv()
    )
  })
---

```{r child="mydefs.Rmd"}
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
plot_cadlag <- function(x,y,blacks=TRUE,whites=TRUE, treshold=0.5, point_size=1, ...) {
  points = tibble(x,y, is_black=FALSE, is_white = FALSE)
  # points$is_black[1]=TRUE
  points$is_black[nrow(points)]=TRUE
  
  i = 1
  while (i < nrow(points)) {
    if (abs(points$y[i+1]-points$y[i])>treshold) {
      x0 = points$x[i]
      y0 = points$y[i+1]
      points$is_white[i]=TRUE
      points = points %>% 
        add_row(x=x0, y=y0, is_black=TRUE, is_white=FALSE, .before=i+1)
      points  
       i=i+2}
    else {
      i=i+1
    }
  }
  lpoints = lead(points) %>% 
    rename_with(
      ~ paste0(.x,"end", recycle0 = TRUE)
    )
  
  lines = bind_cols(points, lpoints) %>%
    mutate( is_jump = is_white) %>% 
    filter( row_number() <= n()-1 ) 

  out = ggplot(lines, aes(x=x, y=y, xend=xend, yend=yend, linetype=is_jump)) +
    geom_segment()+
    scale_linetype_manual(name = "Linetype", values=c(1, 0));
  if (blacks) {
    out = out +  geom_point(data = subset(lines, is_black), aes(x=x,y=y), size=point_size)
  }
  
  if (whites) {
   out = out +  geom_point(data = subset(lines, is_whiteend), aes(x=xend,y=yend), shape=21, size=point_size, fill="white")
  }
  out = out + theme(legend.position="none")
  return(out)
}  
```


# The Poisson Process
<div style="counter-reset: thechapter 9;"> </div>

All stochastic processes we have seen so far were *discrete*, i.e., their time
set was $\N_0=\set{0,1,\dots}$ or its subset. The Poisson process is the first
**continuous-time** process we are going to encounter. That means that the time
set is the set $[0,\infty)$ of all nonnegative real numbers. The general theory
of continuous-time processes is significantly more complex than the theory of
discrete processes, but there are a few examples, like the Poisson process, that
are both simple to describe and analyze and have numerous applications.  


Before we give a rigorous definition of the the Poisson process, let us describe
how it works heuristically. We start with a number $\Delta t>0$ and conside the
sequence 
\begin{align}
[0,\Delta t],\  [\Delta t, 2 \Delta t], \ [2 \Delta t, 3\Delta t],\  \dots
\end{align}
of intervals in $[0,\infty)$ - we think of them as boxes. For each of
these boxes, we toss a biased coin with the probability $p\in (0,1)$ of
obtaining heads ($H$) and place a ball in the box if an only if $H$ is obtained. Here is
a simple simulation of the this procedure; we take $\Delta t = 0.05$, $p=0.2$
and display the contents of the first $40$ boxed, i.e., up to $t=2$:

```{r echo=FALSE, fig.height=0.6, fig.width=10, fig.align = "center"}
library(ggplot2)
library(ggh4x)
set.seed(5224)
p = 0.2
n = 40
T = 4
Dt = T/n
ts = Dt*(1:n)
ts0 = c(0,ts)
boxes = ts[rbinom(n,1,p) == 1]-Dt/2
df = data.frame(boxes)
ggplot(df,aes(x=boxes))+
  geom_dotplot(binwidth = Dt-0.001, dotsize = 0.8, fill = "red", stroke = 0.8) +
  scale_x_continuous(
    minor_breaks = seq(0,T,by = Dt), 
    breaks = seq(0,T,by = 1),
    limits=c(0,T),
    guide = "axis_minor")+
  scale_y_continuous(NULL, breaks = NULL)+
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.title.x = element_blank(),
    axis.line.x = element_line(linewidth = 0.0),
    axis.text.x = element_text(hjust = 1, size = 11), 
    ggh4x.axis.ticks.length.minor = rel(1))
```   

One way to describe the obtained allocation of balls to boxes is through the 
so-called **counting process** $\{N_n\}_{n\in\mathbb N_0}$, where $N_n$ is the total number of balls in the boxes numbered $1,2,\dots, n$ and $N_0=0$. 
The realization of the counting process 
corresponding to the allocation from above looks like this.

```{r echo=FALSE, fig.align = "center"}
library(ggplot2)
library(ggh4x)
has = hist(boxes, breaks=ts0, plot=FALSE)$counts
df = data.frame(n=0:length(has), count = c(0,cumsum(has)))

ggplot(df,aes(x=n, y=count))+
  geom_point()+
  scale_y_continuous(breaks = seq(0,max(df$count)+1,5))+
  theme_minimal()

```   
Note that $N$ is still a discrete process. To make it continuous, we give it 
the time-index $[0,\infty)$ and for $t\geq 0$ define $N_t$ as the number of 
balls in the boxes completely contained in $[0,t]$, i.e., in the first $\lfloor t/\Delta t \rfloor$ boxes. This amounts to taking the trajectory of the discrete process $N$, 
multiplying the times index by $\Delta t$ and keeping its value constant throughout each interval $[k \Delta t, (k+1) \Delta t)$:

```{r, echo=FALSE, warnings=FALSE, message=FALSE, fig.align='center', fig.width=7}
library(tidyverse)
plot_cadlag(x=c(0,ts),y=c(0,cumsum(has))) +
  theme_minimal()+
  scale_y_continuous(breaks = seq(0,sum(has)+1,2))+
  scale_x_continuous(breaks = ts0[c(TRUE,FALSE)], limits=c(0,max(ts)))+
  theme(legend.position="none")
  
```

The black and white points are added to stress the fact that at the value at each discontinuity is equal to the limit from the right.

The Poisson process will appear in the limit the counting processes as $\Delta t
\to 0$. This limit will exist, however, only if we let $p \to 0$ at the same rate.
To see why, we take $t = n \Delta t$ and observe that 
$N_t$ has the binomial distribution 
with parameters $n$ and $p$, and, consequently $\EE[ N_t] = np$. 
Letting $\Delta t \to 0$ while keeping $p$ and $t$ fixed amounts to letting
$n\to \infty$ and, with it, the expected value of $\EE[ N_t]$. The way to keep
this expectation from exploding is to decrease the probability $p$ to $0$ so
that $p/ \Delta t \to \ld$ for some $\ld>0$.
While we will not prove it in these notes (or even explain exactly what it means), the corresponding counting processes will converge towards 
a continuous-time stochastic process, called the  **Poisson
process** with parameter $\ld>0$ which is constructed as follows. We start with a
sequence of  independent random variables with the exponential
distribution with parameter $\ld$ and define the random variables $P_0,P_1,\dots$
by
\begin{align}
  P_0 = 0 \eand  P_n = \sum_{k=1}^n T_k \efor n\in\N.
\end{align}
and then set
\begin{align}
N_t = \max \{ n\in\N_0 \, : \, P_n \leq t\}
\end{align}
Simulation:
```{r}
simulate_Poisson <- function(T_max=5, lambda=1) {
  N_max = lambda * T_max * 10
  T = 0; N=0;
  while (T[length(T)] < T_max) {
    T=rbind(T,T[length(T)]+rexp(1,rate = lambda));
  }
  return(T[2:(length(T)-1)])
}
T=simulate_Poisson(lambda=2)
plot_cadlag(x=T, y=0:(length(T)-1)-1)
```







## Additional Problems for Chapter 9 

<!--
  sl-prob-01
  ------------------------------------------------
-->

<!-- ::: {.exercise} -->

<!-- ```{r child="problems/05_Stationary_Distributions/sl-prob-01_prb.Rmd"} -->
<!-- ``` -->

<!-- ::: -->


<!-- <details> -->
<!-- <summary> Click for Solution </summary> -->
<!-- ::: {.solution} -->
<!-- \ -->
<!-- ```{r child="problems/05_Stationary_Distributions/sl-prob-01_sol.Rmd"} -->
<!-- ``` -->
<!--  -->
<!-- ::: -->
<!-- </details> -->

## Endnotes

